Well-quasi-orderings on word languages \\ main edb3a6bc24be4a12fe4f383c75384a6dff9d1a3c \\ 2025-10-02 17:15:30 +0200

### 2025-10-06 src/infixes.tex
new: answering todos in a stupid way

\begin{corollary}
    \label{bounded-wqo-dwclosed:cor}
    Let $L$ be a \kl{bounded language} of $\Sigma^*$. Then,
    $L$ is a \kl{well-quasi-order} when endowed with the
    \kl{infix relation} if and only if $\dwset[\infleq]{L}$ is.
\end{corollary}
\begin{proof}
    Because $L \subseteq \dwset[\infleq]{L}$, the right-to-left implication
    is trivial.
    For the left-to-right implication, let us assume that $L$ is a
    \kl{well-quasi-ordered} language for the \kl{infix relation}.
    Then $L$ is included in a finite union 
    of products of \kl{chains}: \todo{reference}
    \begin{equation*}
        L \subseteq \bigcup_{i = 1}^n S_i \cdot P_i \quad .
    \end{equation*}
    Remark that the \kl{downwards closure} of a product of two \kl{chains}
    is a finite union of products of two chains. \todo{reference/proof}.
    As a consequence, we conclude that $\dwset[\infleq]{L}$ is itself included
    in a finite union of products of \kl{chains}.
    Note that this also proves that $\dwset[\infleq]{L}$ is a \kl{bounded language},
    hence that it is \kl{well-quasi-ordered} by the \kl{infix relation} 
    via
    \cref{bounded-language:thm}.
\end{proof}

One may think that all \kl{downwards closed} languages for the \kl{infix
relation} that are \kl{well-quasi-ordered} are \kl(language){bounded}. Note
that this is what happens in the case of the \kl{subword embedding}, where any
\kl{downwards closed} language for this relation is characterized by finitely
many excluded \kl{subwords}, hence provides a \kl{bounded language} \todo{wrong? $\{a,b\}^*$ is downward closed but not bounded}. However, this
is not the case for the \kl{infix relation}, as we will now illustrate with the
following two examples.

\begin{example}
    \label{dwclosed-wqo-not-finite-excl:ex}
    Let $L \defined a^* b^* \cup b^* a^*$ \todo{However, this language is bounded, so this seems to be a non sequitur}. This language is \kl{downwards
    closed} for the \kl{infix relation}, is \kl{well-quasi-ordered} for the
    \kl{infix relation}, but is characterized by an \emph{infinite} number 
    of excluded infixes, respectively of the form $ab^ka$ and $ba^kb$ where $k \geq 1$.
\end{example}

>>>>

\begin{corollary}
    \label{bounded-wqo-dwclosed:cor}
    Let $L$ be a \kl{bounded language} of $\Sigma^*$. Then,
    $L$ is a \kl{well-quasi-order} when endowed with the
    \kl{infix relation} if and only if $\dwset[\infleq]{L}$ is.
\end{corollary}
\begin{proof}
    Because $L \subseteq \dwset[\infleq]{L}$, the right-to-left implication
    is trivial.
    For the left-to-right implication, let us assume that $L$ is a
    \kl{well-quasi-ordered} language for the \kl{infix relation}.
    Then $L$ is included in a finite union 
    of products of \kl{chains}:
    \begin{equation*}
        L \subseteq \bigcup_{i = 1}^n S_i \cdot P_i \quad .
    \end{equation*}
    Remark that the \kl{downwards closure} of a product of two \kl{chains}
    is a finite union of products of two chains.
    As a consequence, we conclude that $\dwset[\infleq]{L}$ is itself included
    in a finite union of products of \kl{chains}.
    Note that this also proves that $\dwset[\infleq]{L}$ is a \kl{bounded language},
    hence that it is \kl{well-quasi-ordered} by the \kl{infix relation} 
    via
    \cref{bounded-language:thm}.
\end{proof}

One may think that all \kl{downwards closed} languages for the \kl{infix
relation} that are \kl{well-quasi-ordered} are \kl(language){bounded}. Note
that this is what happens in the case of the \kl{subword embedding}, where any
\kl{downwards closed} language for this relation is characterized by finitely
many excluded \kl{subwords}, hence provides a \kl{bounded language}. However, this
is not the case for the \kl{infix relation}, as we will now illustrate with the
following two examples.

\begin{example}
    \label{dwclosed-wqo-not-finite-excl:ex}
    Let $L \defined a^* b^* \cup b^* a^*$. This language is \kl{downwards
    closed} for the \kl{infix relation}, is \kl{well-quasi-ordered} for the
    \kl{infix relation}, but is characterized by an \emph{infinite} number 
    of excluded infixes, respectively of the form $ab^ka$ and $ba^kb$ where $k \geq 1$.
\end{example}

### 2025-10-06 wqo-on-words.lncs.tex
fix: knowledges + refs + trying to prove uur decidability for pure morphic words
new: answering todos in a stupid way

% links inside the document
\usepackage{hyperref}
\usepackage[capitalise,noabbrev,nameinlink]{cleveref}
\Crefname{assumption}{Assumption}{Assumptions}
\usepackage[composition,hyperref,xcolor,cleveref]{knowledge}
\knowledgeconfigure{notion}

@@

\begin{document}
%
\title{Well-quasi-orderings on word languages \\
       main dd2237175323943c3893f9ae93107aba9f6049a5 \\
       2025-09-08 16:09:49 +0200
}
\titlerunning{WQO on Words}

\author{
        Nathan Lhote\inst{2}\orcidID{0000-0003-3303-5368} \and
        Aliaume Lopez\inst{1}\orcidID{0000-0002-4205-327X} \and
        Lia Schütze\inst{3}\orcidID{0000-0003-4002-5491}}
\authorrunning{N. Lhote and A. Lopez and L. Schütze}
\institute{
INP Bordeaux, LaBRI, CNRS
 \and
Aix-Marseille University
 \and
Max Planck Institute for Software Systems
}

>>>>

% links inside the document
\usepackage{hyperref}
\usepackage[capitalise,noabbrev,nameinlink]{cleveref}
\Crefname{assumption}{Assumption}{Assumptions}
\Crefname{conjecture}{Conjecture}{Conjectures}
\usepackage[composition,hyperref,xcolor,cleveref]{knowledge}
\knowledgeconfigure{notion}

@@

\begin{document}
%
\title{Well-quasi-orderings on word languages \\
       main edb3a6bc24be4a12fe4f383c75384a6dff9d1a3c \\
       2025-10-02 17:15:30 +0200
}
\titlerunning{WQO on Words}

\author{
        Nathan Lhote\inst{3}\orcidID{0000-0003-3303-5368} \and
        Aliaume Lopez\inst{2}\orcidID{0000-0002-4205-327X} \and
        Lia Schütze\inst{1}\orcidID{0000-0003-4002-5491}}
\authorrunning{N. Lhote and A. Lopez and L. Schütze}
\institute{
Max Planck Institute for Software Systems
 \and
INP Bordeaux, LaBRI, CNRS
 \and
Aix-Marseille University
}
